The vertical Fermi and eROSITA bubbles inclined jets: A proof-of-concept study

### 2022-01-06 paper.tex
update

  With the help of the isothermal and hydrostatic equilibrium conditions,\
  and assuming the interfaces between isothermal disk and atmosphere are in pressure equilibrium,\
  we can write the steady-state gaseous density distributions,\
  confined in the total potential, of the disk and the Galactic atmosphere as\
  \begin{subequations}
  \begin{align}
     \displaystyle \rho_{\text{isoDisk}}(z) = \rho_{\text{isoDisk}}^{\text{peak}}
     \exp\left[-\frac{\Phi_{\text{total}}(z)}{k_{B}T_{\text{isoDisk}}/m_{\text{p}}}\right]&\label{isothermal-disk-density}\\
     \text{, if $|z| < z_{0}$}& \nonumber \\
     \nonumber\\
     \displaystyle \rho_{\text{atmp}}(z) = \rho_{\text{atmp}}^{\text{peak}}
     \exp\left[-\frac{\Phi_{\text{total}}(z)}{k_{B}T_{\text{atmp}}/m_{\text{p}}}\right]&\label{isothermal-atmp-density}\\
     \text{, otherwise,}& \nonumber
  \end{align}
  \label{disk-atm-sys}
  \end{subequations}
  where $m_{\text{p}}$ is the proton mass,\
  $T_{\text{isoDisk}}$ and $T_{\text{atmp}}$ is the temperature of the isothermal disk and atmosphere,\
  $\rho_{\text{isoDisk}}^{\text{peak}}$ and $\rho_{\text{atmp}}^{\text{peak}}$ is the peak mass density\
  of the disk and atmosphere on the mid-plane $z=0$.

  We tabulate parameters in the first four\
  categories of \Cref{table-parameters},\
  except for $\rho_{\text{atmp}}^{\text{peak}}$ that\
  can be derived from the other known parameters and pressure equilibrium condition\
  on the interfaces $(z=\pm z_{0})$ between the disk and atmosphere.\

  The density profile of \Cref{disk-atm-sys} is shown in \Cref{fig__initial-density-profile}\
  and compared to the observed result \citep{Miller_2013} beyond 1 kpc.\

  Note that there is an difficulty in disentangling the contribution\
  of the Local Bubble, a supernova remnant in which the Solar System is embedded \citep{Snowden1990},\
  and the contribution from solar wind charge-exchange processes, which produce soft\
  X-ray emission throughout the Solar System. As a result, the density profile of the Galactic halo remains unclear\
  \citep{BlandHawthorn2016}.

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__initial-density-profile.png}
    \caption{The profile of $\rho_{\text{isoDisk}}(z)$ along z-axis.}
    \label{fig__initial-density-profile}
  \end{figure}

  \begin{figure}
      \includegraphics[width=\columnwidth]{figures/fig__numberDensityHistogram.png}
      \caption{The initial density distribution of the number density\
               for a settling ISM without jet source.} % caption
      \label{fig__numberDensityHistogram}
  \end{figure}

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__zoom-in-disk.png}
    \caption{The slice of pressure (top), temperature (middle), and number density (bottom)\
             in the y-z plane passing through the center of the disk.}
    \label{fig__zoom-in-disk}
  \end{figure}

@@

  The density of clumpy disk can thus be obtained by taking the scalar products of\
  $f(\bold{x})$ with $\rho_{\text{isoDisk}}(z)$ over all cells within the disk, i.e.,\
  $\displaystyle\rho_{\text{ismDisk}}(\bold{x}) =\
  f(\bold{x}) \rho_{\text{isoDisk}}(z)$.\
  Also, the pressure equilibrium within the clumpy disk implies that the temperature of disk is
  $\displaystyle T_{\text{ismDisk}}(\bold{x}) =\
  T_{\text{isoDisk}}(z)\rho_{\text{isoDisk}}(z)/\rho_{\text{ismDisk}}(\bold{x})$.

@@

@@

\subsection{Inclined jet injection}
  Early observation \citep{Gallimore2006} has proposed that\
  there is a lack of preferred orientation of jets with respect to the plane of the disc.\
  There are a number of galaxies in which the jets are inclined to the disc normal\
  (e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
  including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).

>>>>

  With the help of the isothermal and hydrostatic equilibrium conditions,\
  and assuming the interfaces between isothermal disk and atmosphere are in thermal pressure equilibrium,\
  we can write the steady-state gaseous density distributions,\
  confined in the total potential, of the disk and the Galactic atmosphere as\
  \begin{subequations}
  \begin{align}
     \displaystyle \rho_{\text{isoDisk}}(z) = \rho_{\text{isoDisk}}^{\text{peak}}
     \exp\left[-\frac{\Phi_{\text{total}}(z)}{k_{B}T_{\text{isoDisk}}/m_{\text{p}}}\right]&\label{isothermal-disk-density}\\
     \text{, if $|z| < z_{0}$}& \nonumber \\
     \nonumber\\
     \displaystyle \rho_{\text{atmp}}(z) = \rho_{\text{atmp}}^{\text{peak}}
     \exp\left[-\frac{\Phi_{\text{total}}(z)}{k_{B}T_{\text{atmp}}/m_{\text{p}}}\right]&\label{isothermal-atmp-density}\\
     \text{, otherwise,}& \nonumber
  \end{align}
  \label{disk-atm-sys}
  \end{subequations}
  where $m_{\text{p}}$ is the proton mass,\
  $T_{\text{isoDisk}}$ and $T_{\text{atmp}}$ is the temperature of the isothermal disk and atmosphere,\
  $\rho_{\text{isoDisk}}^{\text{peak}}$ and $\rho_{\text{atmp}}^{\text{peak}}$ is the peak mass density\
  of the disk and atmosphere on the mid-plane $z=0$.

  We tabulate parameters in the first four\
  categories of \Cref{table-parameters},\
  except for $\rho_{\text{atmp}}^{\text{peak}}$ that\
  can be derived from the other known parameters and thermal pressure equilibrium condition\
  on the interfaces $(z=\pm z_{0})$ between the disk and atmosphere.\

  The density profile of \Cref{disk-atm-sys} is shown in \Cref{fig__initial-density-profile}.\
  Beyond the core radius ($\sim 2 \text{ kpc}$) the gaseous density decreases rapidly as a power-law.

%  The density profile of \Cref{disk-atm-sys} is shown in \Cref{fig__initial-density-profile}\
%  and compared to the observed result \citep{Miller_2013} beyond 1 kpc.\
%
%  Note that there is an difficulty in disentangling the contribution\
%  of the Local Bubble, a supernova remnant in which the Solar System is embedded \citep{Snowden1990},\
%  and the contribution from solar wind charge-exchange processes, which produce soft\
%  X-ray emission throughout the Solar System. As a result, the density profile of the Galactic halo remains unclear\
%  \citep{BlandHawthorn2016}.

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__initial-density-profile.png}
    \caption{The density profile of the isothermal disk (red pluses) and\
             atmosphere (blue crosses) along the positive z-axis.\
             The density distribution is derived from hydrostatic equilibrium,\
             the interface ($z=0.1$) between the isothermal disk\
             and the atmosphere is pressure balanced.}
    \label{fig__initial-density-profile}
  \end{figure}

@@

  The density of clumpy disk can thus be obtained by taking the scalar products of\
  $f(\bold{x})$ with $\rho_{\text{isoDisk}}(z)$ over all cells within the disk, i.e.,\
  $\displaystyle\rho_{\text{ismDisk}}(\bold{x}) =\
  f(\bold{x}) \rho_{\text{isoDisk}}(z)$.\
  Also, the thermal pressure equilibrium within the clumpy disk implies that the temperature of disk is
  $\displaystyle T_{\text{ismDisk}}(\bold{x}) =\
  T_{\text{isoDisk}}(z)\rho_{\text{isoDisk}}(z)/\rho_{\text{ismDisk}}(\bold{x})$.

@@

  \Cref{fig__zoom-in-disk} shows a close-up view of the\
  pressure, temperature, and number density slices
  in the y-z plane through the center of the disk.

  \Cref{fig__numberDensityHistogram} plots the volume filling factor as a function of\
  initial number density within the disk without jet source.\

  \begin{figure}
      \includegraphics[width=\columnwidth]{figures/fig__numberDensityHistogram.png}
    \caption{The volume filling factor as a function of\
             initial number density within the disk without jet source.\
             The regions covered by vertical bands represent the allowable number densities \citep{peak-ism-density} for\
             hot ionized, warm neutral (WNM), warm ionized (WIM), cold nuetral mediums (CNM), and molecular clouds.}
      \label{fig__numberDensityHistogram}
  \end{figure}

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__zoom-in-disk.png}
    \caption{Close-up view of the\
             pressure (top), temperature (middle), and number density (bottom) slices\
             in the y-z plane through the center of the disk.
             }
    \label{fig__zoom-in-disk}
  \end{figure}

@@

\subsection{Inclined jet injection}
  Early observation \citep{Gallimore2006} has proposed that\
  there is a lack of preferred orientation of jets with respect to the plane of the disc.\
  A number of galaxies in which the jets are inclined to the disc normal\
  (e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
  including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).
